---
title: "2020-11-04 World Bank CO2"
author: "Paul Ramírez"
date: "04/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Wrangling

### Loading of raw data (CO2 (kt) RAW DATA)

```{r}
library(tidyverse)
rel_path_excel = "World_Bank_CO2.xlsx"
excel_sheets <- readxl::excel_sheets(rel_path_excel)
show(excel_sheets)
excel_sheets[3]
excel_sheets[7]
raw_CO2_data <- readxl::read_excel(rel_path_excel, sheet = excel_sheets[3])
raw_CO2_per_capita_data <- readxl::read_excel(rel_path_excel, sheet = excel_sheets[7])
class(raw_CO2_data)
head( raw_CO2_data, 10)

```
### Tidying data, `gather` function

Selecting columns names of "year" columns
```{r}

year_columns1 = colnames( raw_CO2_data %>% select(5:ncol(raw_CO2_data)) )

year_columns2 = colnames( raw_CO2_per_capita_data %>% select(5:ncol(raw_CO2_per_capita_data)) )

```

Unpivoting "year" columns to rows using `gather`

```{r}
tidy_CO2_data <- raw_CO2_data %>%
  gather(year_columns1, key = "Year", value = "CO2 emissions (kt)")

tidy_CO2_per_capita_data <- raw_CO2_per_capita_data %>%
  gather(year_columns2, key = "Year", value = "CO2 emissions (metric tons per capita)")

```
Removing unnecessary columns

```{r}
show( colnames( tidy_CO2_data ) )
tidy_CO2_data <- tidy_CO2_data %>%
  select(-`Indicator Name`, -`Indicator Code`)
show( colnames( tidy_CO2_data ) )
tidy_CO2_data %>% arrange(`Country Code`, Year) %>% head(10)

show( colnames( tidy_CO2_per_capita_data ) )
tidy_CO2_per_capita_data <- tidy_CO2_per_capita_data %>%
  select(-`Indicator Name`, -`Indicator Code`)
show( colnames( tidy_CO2_per_capita_data ) )
tidy_CO2_per_capita_data %>% arrange(`Country Code`, Year) %>% head(10)


```
## Joining CO2 dataset with CO2 per capita to obtain the final clean dataset

First we need to create an extra column with the Country Code - Year pairings so they are a unique primary key in both tables

```{r}

tidy_CO2_data <- tidy_CO2_data %>% 
  mutate( CC_Year = str_c(`Country Code`, Year, sep = "_") )

tidy_CO2_per_capita_data <- tidy_CO2_per_capita_data %>% 
  mutate( CC_Year = str_c(`Country Code`, Year, sep = "_") )

tidy_CO2_per_capita_data_select <- tidy_CO2_per_capita_data %>%
  select(-`Country Name`, -`Country Code`, -Year)

tidy_CO2_final <- tidy_CO2_data %>%
  full_join( tidy_CO2_per_capita_data_select, by = "CC_Year" ) %>%
  select(-CC_Year)
tidy_CO2_final %>% 
  arrange(`Country Code`, Year) %>%
  head(10)

```







